Well-quasi-orderings on word languages

### 2024-08-16 src/infixes.tex
adding decidability

    Let $u$ and $v$ be two words in the \kl{gap language}
    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 \leq \ell \leq
    \card{\canrun(r)}$. We have corresponding runs $s_1, s_2 \in R$ and
    embeddings $f_1 \in E(r,s_1)$ and $f_2 \in E(r,s_2)$ such that
    $\GapWord{f_1}{\ell} = u$ and $\GapWord{f_2}{\ell} = v$. Using the
    amalgamation property, we can inductively construct runs $r_k \in R$ and
    embeddings $f_k \in E(r,s_k)$ such that $\GapWord{f_k}{\ell} = u v^k u$,
    and $\GapWord{f_k}{i}$ is obtained by concatenating in some order two
    occurrences of $\GapWord{f_1}{i}$ and $k$ occurrences of
    $\GapWord{f_2}{i}$. Because $L$ is \kl{well-quasi-ordered} for the
    \kl{infix relation}, there exists $i < j$ such that $\yieldrun(r_i) \infleq
    \yieldrun(r_j)$. Assuming that $\card{\canrun(r)}$ is greater than $2$, and
    that we have extracted an increasing sequence of indices so that
    $\alpha(i+1) > \factorial{\card{\yieldrun(r_i)}}$, we conclude that $u
    v^{\alpha(i)} u$ is an \kl{infix} of $\yieldrun(r_j)$.
    Because of the number of repetitions, a \emph{large} infix of $v^{\alpha(i)}$ must be
    an infix of some $(\GapWord{f_2}{p})^{\alpha(j)}$.
    We conclude that $(\GapWord{f_2}{p})^{\alpha(j)}$ and $v$
    have a common period. Refining the analysis, we also conclude
    that $uv^{q}$ and $(\GapWord{f_2}{p})^{\alpha(j)}$ have the same period,
    which proves that $uv^{q}u$ belongs to some $\InfPeriodChain{x}$
    \textbf{todo}.

    \begin{quote}
        Idea: when pumping like this, we can always find
        large blocks of $\GapWord{f_2}{p}$, so this looks like
        something suitable for \cref{pumping-periods:lem}.
        This should allow us to conclude that every pair of words
        $u,v$ in a given gap language
        are infixes of \emph{the same} periodic word repeated.
    \end{quote}

    Then, we can leverage \cref{pumping-periods:lem}
    to conclude that $L$ is a finite union of \kl{chains}
    for the \kl{infix relation}.
\end{proofof}

>>>>

    Let $u$ and $v$ be two words in the \kl{gap language}
    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 < \ell <
    \card{\canrun(r)}$. We have corresponding runs $s_1, s_2 \in R$ and
    embeddings $f_1 \in E(r,s_1)$ and $f_2 \in E(r,s_2)$ such that
    $\GapWord{f_1}{\ell} = u$ and $\GapWord{f_2}{\ell} = v$. Using the
    amalgamation property, we can inductively construct runs $r_k \in R$ and
    embeddings $f_k \in E(r,s_k)$ such that $\GapWord{f_k}{\ell} = u v^k u$,
    and $\GapWord{f_k}{i}$ is obtained by concatenating in some order two
    occurrences of $\GapWord{f_1}{i}$ and $k$ occurrences of
    $\GapWord{f_2}{i}$. 
    This means that $\canrun(r_i)$ has the shape
    \begin{equation*}
        v_0^{x_1} u_0 v_0^{x_2} u_0 v_0^{x_3} \canrun(r_i)_1 
        v_1^{x_4} u_1 v_1^{x_5} u_1 v_1^{x_6} \canrun(r_i)_2 
        \cdots
        v_{\ell}^{x_k} u_{\ell}
        v_{\ell}^{x_{k+1}}  u_{\ell}
        v_{\ell}^{x_{k+2}}
        \canrun(r_i)_{\ell}
        u v^i u
        \canrun(r_i)_{\ell+1}
        \cdots
        \quad .
    \end{equation*}

    Now, leveraging \cref{pumping-periods:lem}, we get that $u v^k u$ is an
    infix of some $x^p$ where $x$ has size smaller or equal than $u$ and $v$.
    This proves  that $u$ and $v$ are comparable for the prefix, infix, and
    suffix relations. In particular, we have proven that $\GapLanguage{r}{i}$
    is bounded, because of \cite[Lemma 4.1]{ASZZ24}, when $0 < i <
    \card{\canrun(r)}$.

    For the edge cases, we do a very similar construction. For $i = 0$, we let
    $\canrun(r_i) = v^k u \canrun(r_1) \dots$, we conclude similarly. For $i =
    \card{\canrun(r)}$, we let $\canrun(r_i) = \dots
    \canrun(r)_{\card{\canrun(r) - 1}} u v^k$.
\end{proofof}

The abstract definition of \kl{amalgamation system} can be enriched by
effectiveness assumptions, as explained in \cite[Section 3.1.4]{ASZZ24}. In
this setting, we conjecture that the following decidability result holds.
